---
title: Final Exam
output: html_document
---

# Final Exam

## Q1 Grade data Analysis
``` {R}
data = read.csv("./GRADES.dat", header=T, sep="")
head(data)
```


### (a) Clustering

Using K-mean to cluster
``` {R}
cl = kmeans(data, 3)
cl
```

And show the first two axis of PCA to evaluate the cluster.
``` {r}
pc_data = princomp(data ,cor=TRUE)
new_data = t(t(pc_data$loadings[, 1:2]) %*% t(as.matrix(data)))
plot(new_data,col=cl$cluster)
points(cl$centers, col=1:2, pch=8, cex=2)
```

And there mean performance in each cluster
```
cl$centers
```

### (b) Factor Analysis

Perform FA

``` {r}
library(psych)
data = read.csv("./GRADES.dat", header=T, sep="")
head(data)
data_cor = head(data)
fa_data = principal(data_cor, nfactors=4, rotate='varimax')
print(fa_data$loading, cutoff=.5)
```

You can see the result.

Lab score has high correlated to HW, exam1 and exam2.

Popquiz is related to exam1 and finalexam. Also in other component, it related to exam2.

And in the scree plot, we can see choosing three factors is enough.

``` {r}
a = colSums(fa_data$loadings ^2)
a$sdev = colSums(fa_data$loadings ^2)
screeplot(a)
abline(h=1, col="red")
```

